Topic modeling on EU and the USA research projects

### 2017-06-23 thesis_report/draft_thesis.tex
Discovering topics on the USA projects

In LDA, our ultimate goal is to find the posterior distribution of latent topic variables after observing training data. However, computing this distribution is intractable, so we’re forced to approximate. One approximation approach is to 
use an optimization method called Variational Bayes. In the original paper it's 
being used a variational Bayes approximation of the posterior distribution;\cite{blei2003latent} alternative inference techniques use Gibbs sampling and
expectation propagation. The goal of variational inference is to approximate an
intractable probability distribution $p$, with a tractable one $q$, in a way
that makes them as close as possible given a similarity measure. In this thesis,
variational inference is used to approximate the posterior distribution of
latent variables. A new distribution over the hidden variables $q(z, \beta)$
(called the variational distribution) is  introduced. This new distribution has
properties such that it can be efficiently computed.\cite{Fox2011ATO}

@@

\begin{center}
\begin{tabular}{l*{6}{c}r}
Framework         & Years & No of docs & No of unique tokens & No of words in the corpus \\
\hline
FP4 & 1994-1997 & 36752 & 28038 & 4346604 \\
FP5 & 1998-2001 & 37674 & 30728 & 5166316 \\
FP6 & 2002-2006 & 54212 & 38476 & 8837298 \\
FP7 & 2007-2013 & 87341 & 48277 & 15673424 \\
H2020 & 2014-2017 & 35581 & 32268 & 7379219 \\
\end{tabular}
\end{center}

@@

@@

@@

>>>>

In LDA, our ultimate goal is to find the posterior distribution of latent topic 
variables after observing training data. However, computing this distribution 
is intractable, so we’re forced to approximate. One approximation approach is to 
use an optimization method called Variational Bayes. In the original paper it's 
being used a variational Bayes approximation of the posterior distribution;
\cite{blei2003latent} alternative inference techniques use Gibbs sampling and
expectation propagation. The goal of variational inference is to approximate an
intractable probability distribution $p$, with a tractable one $q$, in a way
that makes them as close as possible given a similarity measure. In this thesis,
variational inference is used to approximate the posterior distribution of
latent variables. A new distribution over the hidden variables $q(z, \beta)$
(called the variational distribution) is  introduced. This new distribution has
properties such that it can be efficiently computed.\cite{Fox2011ATO}

@@

\begin{center}
\begin{tabular}{l*{6}{c}r}
Framework& Years & No of docs & No of unique tokens & No of words in the corpus \\
\hline
FP4 & 1994-1997 & 36752 & 28038 & 4346604 \\
FP5 & 1998-2001 & 37674 & 30728 & 5166316 \\
FP6 & 2002-2006 & 54212 & 38476 & 8837298 \\
FP7 & 2007-2013 & 87341 & 48277 & 15673424 \\
H2020 & 2014-2017 & 35581 & 32268 & 7379219 \\
\end{tabular}
\end{center}

@@

The National Science Foundation mostly supports fundamental research and
education in the non-medicalin science and engineering. Its medical counterpart
is the National Institutes of Health.\footnote{\url{https://en.wikipedia.org/wik
i/National_Science_Foundation}} With STEM being the top priority in USA, most of
the projects founded by NSF are related to engineering, computer science,
mathematics and information systems, in general. The subjects STEM \& Education,
Software engineering, Information systems and Mathematics, arise as topics in
all these years, from 1994-2017.

While NSF promotes the progress of science, it also enhances the effords to
maintain the national health, prosperity, and welfare. Over the last 22 years,
many projects are related to the environment effects of climate change. Topics
like Environmental Science, Atmospheric \& Ocean sciences, Geology, Climate
change and its impact in the United States emerge during all these years. The
last three years there is a focus on the Green energy production, like
hydropower and wind power. Among with climate change, there is many research
related to Biology and Evolution. A less popular subject that appears during the
years 1994-1997 and 2007-2013 is Molecular, Cellular \& Organismal biology and
Genetics. These projects mostly refer to applied genetics, as well as cell
functionality. Topics include receptor biology, regulation of gene expression,
biomedical engineering, patterns of inheritance and molecular genetics.

Another popular subject that appears the last 22 years is Materials and
Chemistry; the projects related to material research are mostly refer to the
materials' properties and applications in engineering. The last 3 years,
researchers from the USA focus on materials applications and desing in
mechanical engineering, bioengineering and chemical \& transport systems. In
2002-2006 many projects are related to the subject of transport that is
sustainable in the senses of social, environmental and climate impacts (Green
transportation).

From 1994-2006 and the last three years, International Research and conferences
came up as a trending topic. Re- searchers have a great interest to collaborate
with other colleagues from all over the world and share their knowledge. A big
part of the projects funded by the EU that is related to international research,
is also related to Physics, in general, and High energy physics in EU (see
Appendix: Wordclouds of topics of the EU).

NSF shows a great interest in international activities across all NSF supported
disciplines. The goal of these projects is to support high quality research and
education, which could not occur without international collaboration. NSF seeks
to catalyze a higher level of international engagement in the U.S. science and
engineering community by promoting International conferences and university
projects and partnerships. An interesting fact regarding international
partnerships is that the European Commission and the Government of the United
States of America concluded on 17 October 2016 an Implementing Arrangement that
simplifies the cooperation between Horizon 2020 projects and US entities.
European and American researchers will be able to work together more closely on
projects funded under Horizon 2020, the European Union's research and innovation
programme.\footnote{\url{http://ec.europa.eu/research/iscp/index.cfm?pg=usa}}

From 2002 up to day, Economic and Social Sciences emerge as a topic in our
dataset. Researchers show a great interest in identifying policies which help
the communities and they also do lots of Qualitative and Quantitative research
to investigate the size or extent of particular issues or trends in society.

Finally, a topic that used to be popular in the past, in 1994-1997 and
2007-2013, is High energy Physics in USA, which has lost its popularity
nowadays. The last 5 years, there is an interest around STEM and its
applications in manufacturing and engineering.

@@

\begin{center}
\begin{figure}
\includegraphics[width=1.0\textwidth, natwidth=707,natheight=483]
{figs/eu-topic-evolution.png}
\caption{Topics over time in EU}
\end{figure}
\end{center}

@@

\subsection{Comparison of the two regions per framework programme}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP4.png}
\caption{Heatmap of the similarity of the two regions during 1994-1997}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP5.png}
\caption{Heatmap of the similarity of the two regions during 1998-2001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP6.png}
\caption{Heatmap of the similarity of the two regions during 2002-2006}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP7.png}
\caption{Heatmap of the similarity of the two regions during 2007-2013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapH2020.png}
\caption{Heatmap of the similarity of the two regions during 2014-2017}
\end{figure}
